---
title:  "Gov 50 Final Project"
author: "Patrick Pieper"
description: This project investigates the carbon emissions data across various countries since '1751', focusing on identifying and visualizing the differences in emissions. The objective is to pinpoint the highest emitters, compare emissions amongst countries, and highlight the gap between them. 
output:
  distill::distill_article:
    self_contained: false
---


## Project thoughts

I was initially interested in the varying levels of carbon emissions across time. However, when reading the data, I was more intrigued by the difference in countries, as opposed to the difference over time. This led me to a broader investigation aiming to visualize and understand the emissions data better. In doing so, i hope to discover significant insights regarding the emissions spectrum - where countries stand and the gaps that exist between them.


## Data Sources

The primary data source for this project is the CO2 emissions data set retrieved from Datahub, within the scope of a project by the Carbon Dioxide Information Analysis Center (CDIAC). This data set contains historical emissions data - recorded yearly across countries - which allows for a comprehensive analysis observing time and country. The data was accessed using R and relevant packages, such as 'jasonlite'.


## Proposal
The goal is to analyze and visualize the carbon emissions data to draw meaningful comparisons between countries, particularly focusing on the top 10 emitters. The analysis will highlight the emissions difference between adjacent countries on the list, providing a visual representation of the emissions gap. This analysis would open the door for a deeper investigation into the policies and factors contributing to the emissions levels in these countries - potentially also allowing for comparison with related metrics, such as GDP. 

Based on the global emissions data, I anticipate that the key emitters will be predominantly industrialized countries, many of which will be categorized as 'Western'. In addition, BRICS countries should also be up on the list. I also expect the difference in emissions to be more significant the higher a country's emissions are.





```{r}
install.packages("jsonlite", repos="https://cran.rstudio.com/")
library("jsonlite")

json_file <- 'https://datahub.io/core/co2-fossil-by-nation/datapackage.json'
json_data <- fromJSON(paste(readLines(json_file), collapse=""))

# get list of all resources:
print(json_data$resources$name)

# print all tabular data(if exists any)
for(i in 1:length(json_data$resources$datahub$type)){
  if(json_data$resources$datahub$type[i]=='derived/csv'){
    path_to_file = json_data$resources$path[i]
    data <- read.csv(url(path_to_file))
    }
}
```


```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(dplyr)
library(ggplot2)
```
```{r}
unique_countries <- data |>
  distinct(Country)
 
```

```{r}
head(data, 10)
```

SHOWCASE HOW DATA PRESENTED

#total emissions for each country across all years

```{r}
total_emissions <- data |>
  group_by(Country) |>
  summarise(total_emissions = sum(Total, na.rm = TRUE)) |>
  arrange(desc(total_emissions))

```


#table with total emissions by country

```{r}
average_emissions <- mean(total_emissions$total_emissions, na.rm=TRUE)

total_emissions |>
ggplot(aes(x = reorder(Country, -total_emissions), y = total_emissions)) +
  geom_col() +
  geom_hline(yintercept = average_emissions) +
  labs(
    x="Country",
    y="Total Emissions",
    title="Total Emissions by Country"
  )
```


#filtered data and averages

```{r}
top10_emissions <- total_emissions |>
  arrange(-total_emissions) |>
  slice(c(1:10))
        top10_emissions
        
average_emissions_top10 <- mean(top10_emissions$total_emissions)
average_emissions_top10
average_emissions_total <- mean(total_emissions$total_emissions)
average_emissions_total
```


#Bar Chart 

```{r}
top10_emissions |>
  ggplot(aes(x = reorder(Country, -total_emissions), y = total_emissions)) +
  geom_bar(stat = "identity") +
  geom_hline(aes(yintercept = average_emissions_total, linetype = "Average for All Countries", color="indianred")) +
  geom_hline(aes(yintercept = average_emissions_top10, linetype = "Average for Top 10 Countries", color="dodgerblue")) +
  labs(x="Country", y="Total Emissions", title="Top 10 Total Emissions by Country") +
  scale_linetype_manual(values=c("Average for All Countries" = "dashed", "Average for Top 10 Countries" = "dotted"), name="Average Line") +
  guides(color = FALSE) + #Remove color legend
  theme(axis.text.x = element_text(angle=45, hjust=1, size=6))
```

The graph showcases the vast difference in emissions among the top 10 countries.The United States - in total over the last ~280 years - has produced more than twice as many emissions than China, in 2nd place. The significant disparity in emissions among these large-scale emitters and the descending trend in emissions values for subsequent countries seems to confirm my assumption that there exists a positive correlation between the amount of emissions and the difference in emissions to a lower country. This trend would indicate that countries producing little emissions have similar data to their neighbors - referring to an arranged order according to emissions.

Comparing the average emissions of these top 10 players to the global average can offer insight into the overall contribution of these 10 to global emissions as well. Using these two approaches, one may substantiate claims that despite being a small number of countries, top emitters are responsible for a disproportionate amount of global emissions - this being more true the closer a country is to the "highest emitter position".

#emission difference


NOTES: If available, I will compare the emissions data to GDP growth - specifically looking at top emitters.


#GDP

```{r}
df<-read.csv("DP_LIVE_16112023222414417.csv")

```

```{r}
unique_location <- df |>
  distinct(LOCATION)
```

#total GDP

```{r}
total_gdp <- df |>
  group_by(LOCATION) |>
  summarise(total_gdp = sum(Value, na.rm = TRUE)) |>
  arrange(desc(total_gdp))

```


#table with total GDP by country

```{r}
average_gdp <- mean(total_gdp$total_gdp, na.rm=TRUE)

total_gdp |>
ggplot(aes(x = reorder(LOCATION, -total_gdp), y = total_gdp)) +
  geom_col() +
  geom_hline(yintercept = average_gdp) +
  labs(
    x="Country",
    y="Total GDP",
    title="Total GDP by Country"
  )
```

#same 10 countries total gdp

```{r}
top_10_emissions_gdp <- total_gdp |>
  filter(LOCATION %in% c("USA", "CHN", "RUS", "GBR", "JPN", "DEU", "IND", "FRA", "CAN"))
top_10_emissions_gdp

average_gdp_top_10_emissions_gdp <- mean(top_10_emissions_gdp$total_gdp)
average_gdp_top_10_emissions_gdp
average_gdp_total <- mean(total_gdp$total_gdp)
average_gdp_total
```

```{r}
top_10_emissions_gdp |>
  ggplot(aes(x = reorder(LOCATION, -total_gdp), y = total_gdp)) +
  geom_bar(stat = "identity") +
  geom_hline(aes(yintercept = average_gdp_total, linetype = "Average for All Countries", color="indianred")) +
  geom_hline(aes(yintercept = average_gdp_top_10_emissions_gdp, linetype = "Average for Top 10 Countries", color="dodgerblue")) +
  labs(x="Country", y="Total GDP", title="Top 10 Total Emission Countries GDP by Country") +
  scale_linetype_manual(values=c("Average for All Countries" = "dashed", "Average for Top 10 Countries" = "dotted"), name="Average Line") +
  guides(color = FALSE) + #Remove color legend
  theme(axis.text.x = element_text(angle=45, hjust=1, size=6))
```

#1992-2014 total gdp and total emissions (solve USSR problem and data only available until 2014)

#filter emissions, aggregate data, identify top 9

```{r}
emissions_filtered <- data |>
  filter(Year >= 1992, Year <= 2014)


total_emissions_1992_2014 <- emissions_filtered |>
  group_by(Country) |>
  summarise(total_emissions = sum(Total, na.rm = TRUE)) |>
  arrange(desc(total_emissions))


top9_emitters <- head(total_emissions_1992_2014, 9)
top9_emitters
```

#filter and aggregate GDP
```{r}
unique_countries_gdp <- df |>
  distinct(LOCATION) |>
  arrange(LOCATION)
```


```{r}
gdp_filtered <- df |>
  filter(TIME >= 1992, TIME <= 2014, LOCATION %in% c("CHN", "USA", "RUS", "IND", "JPN", "DEU", "GBR", "CAN", "KOR"))


total_gdp_1992_2014 <- gdp_filtered |>
  group_by(LOCATION) |>
  summarise(total_gdp = sum(Value, na.rm = TRUE)) |>
  arrange(- total_gdp)

```

# Merged Data

```{r}
emissions_gdp_data <- data.frame(
  Country = c("China", "USA", "Russia", "India", "Japan", "Germany", "UK", "Canada", "Korea"),
  Total_Emissions = c(35489190, 33931362, 10466421, 7964623, 7524624, 5096108, 3249675, 3242871, 2890885),
  Total_GDP = c(161277918, 271880877, 41382019, 75717911, 88956166, 60368485, 42279981, 25127316, 25760099)
)
emissions_gdp_data
```

#linear regression analysis

```{r}
model <- lm(Total_GDP ~ Total_Emissions, data = emissions_gdp_data)
summary(model)
```

```{r}
emissions_gdp_data |>
  ggplot(aes(x = Total_Emissions, y = Total_GDP)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(title = "Relationship between Total Emissions and Total GDP", x = "Total Emissions in metric tons)", y = "Total GDP (in USD)")
```

The positive slope of the regression line indicates a positive correlation between Total Emissions and Total GDP for the top 9 countries by emissions. This confirms my assumption that countries producing greater volumes of emissions have a higher GDP growth than countries with lower emissions. The confidence interval - represented by the shaded line - appears to be narrow enough/confirm the indications of the regression line, although this must be confirmed with statistical reasoning, such as the p-value and R-squared value.












